---
title: "decontX"
author: "Connor H Knight & Faraz Khan"
date: "11/08/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing & Preparing IBRAP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done', position = c('bottom', 'right'))
```

## Installing DecontX

DecontX is contained within the celda package which can be installed using the following function:

```{r, eval=F, echo=T}

BiocManager::install('celda')

```

If celda fails to install it may be due to serveral reasons. A couple of common issues (normally seen in a HPC environment) and solutions can be seen in the [HPC tutorial](https://connorhknight.github.io/IBRAP/articles/HPC_issues.html)

The function itself is typically applied after scrublet (however these are interchangeable) on the raw counts matrix. The workflow then continues on as normal. 

```{r, eval=F, echo=T}

bmmc <- perform.decontX(object = bmmc)

```