---
title: "Installing & Preparing IBRAP"
author: "Connor H Knight"
date: "11/08/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing & Preparing IBRAP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done', position = c('bottom', 'right'))
```

### Installing IBRAP 

IBRAP depends on more than 100 other packages, thus the installation process may take some time and some manual intervention 

Firstly, some Bioconductor packages must be installed prior to the github installation: 

```{r, eval=F, echo=T}

if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install(c("celda", "slingshot", "scran", "scuttle", "SC3", "SingleCellExperiment", "PCAtools", "topGO","shinydashboard", "destiny"))

```

Next, we install IBRAP and other dependencies in a single command:

```{r, eval=F, echo=T}

install.packages('devtools')
devtools::install_github('connorhknight/IBRAP')

```

IBRAP not only uses packages within R, but also through python. For this, we use reticulate to interface python into the R environment. To setup reticulate we must start by installing miniconda:

```{r, eval=F, echo=T}

reticulate::install_miniconda()

```

Once miniconda is installed it will produce an environment called r-reticulate. Next, we must assure the correct packages are installed within the new environment:

```{r, eval=F, echo=T}

prepare.reticulate()

```

Once this has been completed, you are ready to use IBRAP! For a basic understand of a single sample workflow please use the 'getting started' tutorial. For multi sample analysis, you can divert to the integration tutorial. 

Good luck!














