---
title: "Installing & Preparing IBRAP"
author: "Connor H Knight & Faraz Khan"
date: "11/08/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing & Preparing IBRAP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done', position = c('bottom', 'right'))
```

## IBRAP requirements

IBRAP requires R version >=4.0.0 to function.

## Installing IBRAP

IBRAP installation and its dependencies is fully automated and can be performed using the following command:

```{r, eval=F, echo=T}
install.packages('devtools')
devtools::install_github('connorhknight/IBRAP')

```

Once installed we must prepare our reticulate environment which enables us to access python. 
## Preparing reticulate

IBRAP not only uses packages within R, but also through python. For this, we use reticulate to interface python into the R environment. To setup reticulate we must start by installing miniconda:

```{r, eval=F, echo=T}
reticulate::install_miniconda()

```

Once miniconda is installed it will produce an environment called r-reticulate. Next, we must assure the correct packages are installed within the new environment:

```{r, eval=F, echo=T}
prepare.reticulate()

```

Once this has been completed, you are ready to use IBRAP! For a basic understand of a single sample workflow please use the [getting started](https://connorhknight.github.io/IBRAP/articles/using_IBRAP.html) tutorial. For multi-sample analysis, you can divert to the [integration](https://connorhknight.github.io/IBRAP/articles/integration.html) tutorial. 

Good luck!
