---
title: "Installing & Preparing IBRAP"
author: "Connor H Knight"
date: "11/08/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing & Preparing IBRAP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done', position = c('bottom', 'right'))
```

### Installing IBRAP 

IBRAP depends on more than 100 other packages, thus the installation process may take some time and some manual intervention 

Firstly, some Bioconductor packages must be installed: 

```{r, eval=F, echo=T}

if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install(c("celda", "slingshot", "scran", "scuttle", "SC3", "SingleCellExperiment", "PCAtools", "topGO","shinydashboard"))

```

These are packages that are not automatically installed using the subsequent command. 

Next, we install IBRAP and its dependencies:

```{r, eval=F, echo=T}

install.packages('devtools')
devtools::install_github('connorhknight/IBRAP')

```

Once installed, you can enjoy a toolbox of the most commonly used scRNA-seq components, fully integrated. 

